Mean temperature, fluctuation magnitude, and level of biological organization affect biological responses amidst thermal variability: a meta analysis

### 2021-04-29 writing /thesis_draft/_book/thesis.tex
all formatting is in, just need to read for grammar

% To pass between YAML and LaTeX the dollar signs are added by CII
\title{Thermal variability affects biological responses as a result of interactions between mean temperature, fluctuation and biological organization: A Meta Analysis}
\author{Margaret Anne Slein}
% The month and year that you submit your FINAL draft TO THE LIBRARY (May or December)
\date{May 2021}
\division{Mathematics and Natural Sciences}
\advisor{Sam Fey}
\institution{Reed College}
\degree{Bachelor of Arts}
%If you have two advisors for some reason, you can use the following
% Uncommented out by CII
% End of CII addition

@@

In addition to the importance of environmental variability across geologic timescales, organisms also interact with environmental variability on several shorter timescales. Organisms and the ecosystems they inhabit have experienced variation on many scales, from diurnal changes in light availability, to seasonal changes in temperature, to decadal changes in weather patterns (i.e.~El Nino Southern Oscillation). As such, the scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the tenacity and dynamics of ecosystems (khelifa\_usefulness\_2019; Stewart et al., 2013). How variability impacts populations may impact biodiversity, such that only certain species may survive, and how that biodiversity is maintained are key components to the full picture of ecosystem dynamics. As environmental patterns continue to change, from increased mean temperatures to more severe weather events (i.e.~hurricanes), understanding the limits to organisms' ability to cope with changes in environmental conditions is increasingly important for both individual level responses as well as how individual level impacts scale across population and community level responses (Bernardo, 2014; Helmuth et al., 2010; Huey et al., 2012).

@@

\caption{\label{tab:unnamed-chunk-9}Covariates included in models}
\centering
\begin{tabular}[t]{l>{\raggedright\arraybackslash}p{8cm}}
\toprule
\textbf{Variable} & \textbf{Description}\\
\midrule
\cellcolor{gray!6}{mean\_temp\_constant} & \cellcolor{gray!6}{Reported mean temperature in study}\\
flux\_range & Reported range (difference between minimum 
                                     and maximum) of temperature fluctuation experienced\\
org\_level & Level of biological organization, categorized
                                     as either 0 for individual or 1 for population level
\cellcolor{gray!6}{                                     responses}\\
size & Size estimate of study organism, 0 for extra-small
                                     (microscopic), 1 for small (macroscopic), 2 for medium 
                                     (less than a foot), and 3 for large (more than a foot)\\
exp\_age & Life stage at which experiment was conducted, 
                                     categorized as 0 for larval, 1 for juvenile, and 
\cellcolor{gray!6}{                                     2 for adult}\\
\addlinespace
resp\_type & Response variables units classified as rates to traits, 
                                     such that rates included any response variable units with a
                                     “per day” clarifier were classified as rates, all else as
                                     traits\\
stressful & Whether or not the organisms experienced thermally stressful 
                                     temperatures throughout study duration, if temperature range 
                                     exceeded Topt, classified as thermally stressful, else not
                                     thermally stressful. See Appendix Table 1 for references in which
\cellcolor{gray!6}{                                     organisms’ thermal stress metrics were included}\\
\bottomrule
\end{tabular}
\end{table}
\clearpage

@@

\caption[Forest plot of effect sizes]{Forest plot of effects sizes by study and experiment within a study. Indicates that both within and across studies, there were a wide range of effect sizes, likely due to experimental design (including fluctuation range, mean temperature, etc.) Each point represents an average of SDM for multiple responses within each experiment in a single study. Accordingly, the error bars represent an average of variance across multiple responses within a single experiment.}\label{fig:unnamed-chunk-18}
\end{figure}
Overall, even our simple model demonstrates that the intercept is not significant but there was significant heterogeneity across studies (z= 0.79, p \textless{} 0.0001), as demonstrated by Figure 3.1. These results allow us to reject the null hypothesis that there are no significant differences between SMD across the studies included in analysis. As was referenced in the methods, this allowed us to proceed in further detail how different covariates were influencing the effect sizes specifically.

Our full model, including 7 covariates (all and an additional interaction term between fluctuation range and mean temperature, excluding thermal stress), responses were still significantly different from each other when accounting for the nested structure of responses, experiments, and studies (QM=191.3722, df=7, p \textless{} 0.0001). Further, fluctuation range and the interaction term between fluctuation range and mean temperature were statistically significant in the model (df = 139, p \textless{} 0.00001). Organization level was also important in the model, as a statistically significant covariate (df= 139, p \textless{} 0.05).
\begin{figure}

@@

\caption[Coefficient estimates for full model]{Coefficients estimates for each covariate in the full model, where asterix symbols signify statistically significant model predictors.}\label{fig:unnamed-chunk-19}
\end{figure}
\clearpage

Fluctuation range had a positive estimate while the interaction term between range and mean temperature was only slightly negative, both of which were statistically significant at the p \textless{} 0.001 level (Figure 3.2).The most negative estimate (organization level) was significantly influential in our model at a significance level of p \textless{} 0.05. Surprisingly, the most negative estimates of our model coefficients were not statistically significant. In our analysis of thermal stress via a separate random effects model, we found thermal stress was not statistically significant in explaining the effect size (Table 3.1).
\begin{table}[!h]

@@

The interaction between fluctuation range and mean temperature as well as fluctuation range is best displayed in Figure 3.3. At lower mean temperatures, higher fluctuation ranges have a more positive effect on\ldots{}, though that trend starkly ends at about \(24^{\circ}\)C. Beyond \(24^{\circ}\)C, the effect size decreases to nearly half the size of colder temperatures and hovers just above an effect size of one (Figure 3.3).
\begin{figure}

@@

\caption[Scatterplot of relationship between range, mean, and effect size]{Scatterplot describing the relationship between mean temperature, temperature range, and effect size. Temperature fluctuation range from 0-20C, Mean temperature range from 7-33OC. Effect size ranged from -4 to 6, restricted based on the distribution of effect sizes to minimize impact of outliers (see Appendix Figure S2.)}\label{fig:unnamed-chunk-21}
\end{figure}
The trend between organization level and temperature fluctuation range, suggests that populations perform better at higher fluctuation ranges than organisms (Figure 3.4). Interestingly, there were not any significant differences between thermal stress (Figure 3.5), life history metric (Figure 3.6), body size (Figure 3.7), or age (Figure 3.8).

As noted in the methods, when we excluded one highly influential study and included thermal stress (exclusion model), size became much more important in the thermal stress model (df=123, p \textless{} 0.001) as did mean temperature (df=123, p \textless{} 0.001). Surprisingly, organization level, fluctuation range, and the interaction between fluctuation range and mean temperature became insignificant in the thermal stress model (df=123, p \textless{} 0.1).
\begin{figure}

@@

\caption[Effect sizes across temperature range by organization level]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by organization level: individual or population level responses reported in studies.}\label{fig:unnamed-chunk-22}
\end{figure}
\begin{figure}

@@

\caption[Effect sizes across temperature range by thermal stress]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by thermal stress.}\label{fig:unnamed-chunk-23}
\end{figure}
\begin{figure}

@@

\caption[Effect sizes across temperature range by response type]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by life history metric: responses categorized as rate or trait.}\label{fig:unnamed-chunk-24}
\end{figure}
\begin{figure}

@@

\caption[Effect sizes across temperature range by body size]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by body size: extra-small, small, medium, or large.}\label{fig:unnamed-chunk-25}
\end{figure}
\begin{figure}

@@

\caption[Effect sizes across temperature range by experimental age]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20 C, colored by experimental age: larval, juvenile, or adult.}\label{fig:unnamed-chunk-26}
\end{figure}
\hypertarget{discussion}{%
\chapter{Discussion}\label{discussion}}

@@

\caption[Thermal stress data by species]{\label{tab:unnamed-chunk-31}TPC metrics extracted from additional papers to collect thermal history data for each species in the meta analysis (missing data on species X. arvicola, used in Garcia-Ruiz et al 2011)}
\centering
\begin{tabular}[t]{ll>{\raggedright\arraybackslash}p{6cm}}
\toprule
\textbf{Species} & \textbf{Topt} & \textbf{Reference}\\
\midrule
\cellcolor{gray!6}{Drosophila melanogaster} & \cellcolor{gray!6}{24-26.6 C} & \cellcolor{gray!6}{Klepsatel 2013, David 1988}\\
Leptopilina boulardi & 25 C & De lava et al 2016\\
\cellcolor{gray!6}{Mauremys reevesii} & \cellcolor{gray!6}{37.97 +/- 0.64} & \cellcolor{gray!6}{Dang et al 2019}\\
Gryllus firmus & < 30 C & Singh et al 2020\\
\cellcolor{gray!6}{Limnodynastes peronii} & \cellcolor{gray!6}{30 C} & \cellcolor{gray!6}{Seebacher et al 2014}\\
\addlinespace
Limnodynastes tasmaniensis & 30-33 C & Whitehead et al 1989\\
\cellcolor{gray!6}{Platyplectrum ornatum} & \cellcolor{gray!6}{33 C} & \cellcolor{gray!6}{Kern et al 2014}\\
Natrix natrix & </= 30 C & Issac 1997\\
\cellcolor{gray!6}{Coturnix japoncia} & \cellcolor{gray!6}{26 C} & \cellcolor{gray!6}{Alagawany et al 2017}\\
Plestiodon chinensis & 33 C & Baojun et al 2010\\
\addlinespace
\cellcolor{gray!6}{Rhodnis prolixus} & \cellcolor{gray!6}{30-35 C} & \cellcolor{gray!6}{Fresquet and Lazzari 2011}\\
Escherichia coli & 37 C, 35-36 C & Doyle and Schoeni 1984, Bronikowski et al 2001\\
\cellcolor{gray!6}{Salmonella} & \cellcolor{gray!6}{35-36 C} & \cellcolor{gray!6}{Bronikowski et al 2001}\\
Caiman latirostris & 32-33 C & Simoncini et al 2019, Parachu-Marco et al 2017, Pina et al 2002\\
\cellcolor{gray!6}{Trachemys scripta} & \cellcolor{gray!6}{32-33 C} & \cellcolor{gray!6}{Dang et al 2019}\\
\addlinespace
Esteya vermicola & 26 C & Wang et al 2015\\
\bottomrule
\end{tabular}
\end{table}
\backmatter

>>>>

% To pass between YAML and LaTeX the dollar signs are added by CII
\title{Mean temperature, fluctuation magnitude, and level of biological organization affect biological responses amidst thermal variability: a meta analysis}
\author{Margaret Anne Slein}
% The month and year that you submit your FINAL draft TO THE LIBRARY (May or December)
\date{May 2021}
\division{Mathematics and Natural Sciences}
\advisor{Sam Fey}
\institution{Reed College}
\degree{Bachelor of Arts}
%If you have two advisors for some reason, you can use the following
% Uncommented out by CII
% End of CII addition

@@

In addition to the importance of environmental variability across geologic timescales, organisms also interact with environmental variability on several shorter timescales. Organisms and the ecosystems they inhabit have experienced variation on many scales, from diurnal changes in light availability, to seasonal changes in temperature, to decadal changes in weather patterns (i.e.~El Niño Southern Oscillation). As such, the scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the tenacity and dynamics of ecosystems (khelifa\_usefulness\_2019; Stewart et al., 2013). How variability impacts populations may impact biodiversity, such that only certain species may survive, and how that biodiversity is maintained are key components to the full picture of ecosystem dynamics. As environmental patterns continue to change, from increased mean temperatures to more severe weather events (i.e.~hurricanes), understanding the limits to organisms' ability to cope with changes in environmental conditions is increasingly important for both individual level responses as well as how individual level impacts scale across population and community level responses (Bernardo, 2014; Helmuth et al., 2010; Huey et al., 2012).

@@

\caption{\label{tab:unnamed-chunk-9}Covariates included in models}
\centering
\begin{tabular}[t]{l>{\raggedright\arraybackslash}p{8cm}}
\toprule
\textbf{Variable} & \textbf{Description}\\
\midrule
\cellcolor{gray!6}{mean\_temp\_constant} & \cellcolor{gray!6}{Reported mean temperature (°C) in study}\\
flux\_range & Reported range (difference between minimum 
                                     and maximum) of temperature fluctuation (°C) experienced\\
org\_level & Level of biological organization, categorized
                                     as either 0 for individual or 1 for population level
\cellcolor{gray!6}{                                     responses}\\
size & Size estimate of study organism, 0 for extra-small
                                     (microscopic), 1 for small (macroscopic), 2 for medium 
                                     (less than a foot), and 3 for large (more than a foot)\\
exp\_age & Life stage at which experiment was conducted, 
                                     categorized as 0 for larval, 1 for juvenile, and 
\cellcolor{gray!6}{                                     2 for adult}\\
\addlinespace
resp\_type & Response variables units classified as rates to traits, 
                                     such that rates included any response variable units with a
                                     “per day” clarifier were classified as rates, all else as
                                     traits\\
stressful & Whether or not the organisms experienced thermally stressful 
                                     temperatures throughout study duration, if temperature range 
                                     exceeded Topt, classified as thermally stressful, else not
                                     thermally stressful. See Appendix Table 1 for references in which
\cellcolor{gray!6}{                                     organisms’ thermal stress metrics were included}\\
\bottomrule
\end{tabular}
\end{table}
\clearpage

@@

\caption[Forest plot of effect sizes]{Forest plot of effects sizes by study and experiment within a study. Indicates that both within and across studies, there were a wide range of effect sizes, likely due to experimental design (including fluctuation range, mean temperature, etc.) Each point represents an average of SDM for multiple responses within each experiment in a single study. Accordingly, the error bars represent an average of variance across multiple responses within a single experiment.}\label{fig:unnamed-chunk-18}
\end{figure}
Overall, even our simple model (n=140) demonstrates that the intercept is not significant but there was significant heterogeneity across studies (z= 0.79, p \textless{} 0.0001), as demonstrated by Figure 3.1. These results allow us to reject the null hypothesis that there are no significant differences between SMD across the studies included in analysis. As was referenced in the methods, this allowed us to proceed in further detail how different covariates were influencing the effect sizes specifically.

Our full model, including 7 covariates (all and an additional interaction term between fluctuation range and mean temperature, excluding thermal stress), responses were still significantly different from each other when accounting for the nested structure of responses, experiments, and studies (QM = 191.3722, df = 7, p \textless{} 0.0001). Further, fluctuation range (n=140) and the interaction term between fluctuation range and mean temperature (n=140) were statistically significant in the model (df = 7, p \textless{} 0.00001). Organization level (n=140) was also important in the model, as a statistically significant covariate (df= 7, p \textless{} 0.05).
\begin{figure}

@@

\caption[Coefficient estimates for full model]{Coefficients estimates for each covariate in the full model, where asterix symbols signify statistically significant model predictors. Orange error bars represent 95 percent confidence intervals}\label{fig:unnamed-chunk-19}
\end{figure}
\clearpage

Fluctuation range (n=140) had a positive estimate while the interaction term between range and mean temperature was only slightly negative, both of which were statistically significant at the p \textless{} 0.001 level (Figure 3.2).The most negative estimate, organization level (n=140), was significantly influential in our model at a significance level of p \textless{} 0.05. Surprisingly, the remaining negative estimates of our model coefficients, experimental age (n=140) and size (n=140), were not statistically significant. In our analysis of thermal stress via a separate random effects model, we found thermal stress (n=132) was not statistically significant (df=1, p =0.4855) in explaining the effect size (Table 3.1).
\begin{table}[!h]

@@

The interaction between fluctuation range and mean temperature as well as fluctuation range is best displayed in Figure 3.3. At lower mean temperatures, higher fluctuation ranges have a more positive effect on organism responses and performance, though that trend starkly ends at about \(24^{\circ}\)C. Beyond \(24^{\circ}\)C, the effect size decreases to nearly half the size of colder temperatures and hovers just above an effect size of one (Figure 3.3).
\begin{figure}

@@

\caption[Scatterplot of relationship between range, mean, and effect size]{Scatterplot describing the relationship between mean temperature, temperature range, and effect size. Temperature fluctuation range from 0-20°C, Mean temperature range from 7-33°C. Effect size ranged from -4 to 6, restricted based on the distribution of effect sizes to minimize impact of outliers (see Appendix Figure S2.)}\label{fig:unnamed-chunk-21}
\end{figure}
The trend between organization level and temperature fluctuation range, suggests that populations perform better at higher fluctuation ranges than organisms (Figure 3.4). Interestingly, there were not any significant differences between thermal stress (Figure 3.5), life history metric (Figure 3.6), body size (Figure 3.7), or age (Figure 3.8).

As noted in the methods, when we excluded one highly influential study and included thermal stress in the exclusion model (n=132), size became much more important in the exclusion model (df=8, p \textless{} 0.001) as did mean temperature (df=8, p \textless{} 0.001). Surprisingly, organization level, fluctuation range, and the interaction between fluctuation range and mean temperature became insignificant in the exclusion model (df=8, p \textless{} 0.1).
\begin{figure}

@@

\caption[Effect sizes across temperature range by organization level]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by organization level: individual or population level responses reported in studies.}\label{fig:unnamed-chunk-22}
\end{figure}
\begin{figure}

@@

\caption[Effect sizes across temperature range by thermal stress]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by thermal stress.}\label{fig:unnamed-chunk-23}
\end{figure}
\begin{figure}

@@

\caption[Effect sizes across temperature range by response type]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by life history metric: responses categorized as rate or trait.}\label{fig:unnamed-chunk-24}
\end{figure}
\begin{figure}

@@

\caption[Effect sizes across temperature range by body size]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by body size: extra-small, small, medium, or large.}\label{fig:unnamed-chunk-25}
\end{figure}
\begin{figure}

@@

\caption[Effect sizes across temperature range by experimental age]{Linear regression of SMD across temperature fluctuation ranges from 0 to 20°C, colored by experimental age: larval, juvenile, or adult.}\label{fig:unnamed-chunk-26}
\end{figure}
\hypertarget{discussion}{%
\chapter{Discussion}\label{discussion}}

@@

\caption[Thermal stress data by species]{\label{tab:unnamed-chunk-31}TPC metrics extracted from additional papers to collect thermal history data for each species in the meta analysis (missing data on species X. arvicola, used in Garcia-Ruiz et al 2011)}
\centering
\begin{tabular}[t]{ll>{\raggedright\arraybackslash}p{6cm}}
\toprule
\textbf{Species} & \textbf{Topt (°C)} & \textbf{Reference}\\
\midrule
\cellcolor{gray!6}{Drosophila melanogaster} & \cellcolor{gray!6}{24-26.6°C} & \cellcolor{gray!6}{Klepsatel 2013, David 1988}\\
Leptopilina boulardi & 25°C & De lava et al 2016\\
\cellcolor{gray!6}{Mauremys reevesii} & \cellcolor{gray!6}{37.97 ± 0.64 °C} & \cellcolor{gray!6}{Dang et al 2019}\\
Gryllus firmus & ~30°C & Singh et al 2020\\
\cellcolor{gray!6}{Limnodynastes peronii} & \cellcolor{gray!6}{30°C} & \cellcolor{gray!6}{Seebacher et al 2014}\\
\addlinespace
Limnodynastes tasmaniensis & 30-33°C & Whitehead et al 1989\\
\cellcolor{gray!6}{Platyplectrum ornatum} & \cellcolor{gray!6}{33°C} & \cellcolor{gray!6}{Kern et al 2014}\\
Natrix natrix & ~30°C & Issac 1997\\
\cellcolor{gray!6}{Coturnix japoncia} & \cellcolor{gray!6}{26°C} & \cellcolor{gray!6}{Alagawany et al 2017}\\
Plestiodon chinensis & 33°C & Baojun et al 2010\\
\addlinespace
\cellcolor{gray!6}{Rhodnis prolixus} & \cellcolor{gray!6}{30-35°C} & \cellcolor{gray!6}{Fresquet and Lazzari 2011}\\
Escherichia coli & 37°C, 35-36°C & Doyle and Schoeni 1984, Bronikowski et al 2001\\
\cellcolor{gray!6}{Salmonella} & \cellcolor{gray!6}{35-36°C} & \cellcolor{gray!6}{Bronikowski et al 2001}\\
Caiman latirostris & 32-33°C & Simoncini et al 2019, Parachu-Marco et al 2017, Pina et al 2002\\
\cellcolor{gray!6}{Trachemys scripta} & \cellcolor{gray!6}{32-33°C} & \cellcolor{gray!6}{Dang et al 2019}\\
\addlinespace
Esteya vermicola & 26°C & Wang et al 2015\\
\bottomrule
\end{tabular}
\end{table}
\backmatter

### 2021-04-29 writing /thesis_draft/index.Rmd
all formatting is in, just need to read for grammar

---
title: 'Thermal variability affects biological responses as a result of interactions between mean temperature, fluctuation and biological organization: A Meta Analysis'
author: 'Margaret Anne Slein'
date: 'May 2021'
institution: 'Reed College'
division: 'Mathematics and Natural Sciences'
advisor: 'Sam Fey'
department: 'Biology'
degree: 'Bachelor of Arts'
header-includes:
    - \usepackage{setspace}\onehalfspacing
knit: bookdown::render_book
site: bookdown::bookdown_site
params:
  'Install needed packages for {thesisdown}': True
output:
  thesisdown::thesis_pdf: default 
abstract: 'Ecosystems and organisms have experienced variation on many temporal scales, from diurnal changes in light availability, to seasonal changes in temperature, to decadal changes in weather patterns. Though scientific literature has continued to highlight the importance of ecologically relevant studies for understanding the tenacity and dynamics of ecosystems, there has yet to be a quantitative review of how variability impacts biological responses across taxa and scales of biological organization. Here, we present a quantitative meta analysis of how thermal variability impacts biological responses across a variety of taxa, organism sizes, and biological organization. Our results suggest that the range of temperatures an organism experiences is the most important driver of response magnitude, with the interaction of mean temperature and fluctuation range emerging equally as significant predictors in our model. Further, our results also suggest that level of biological organization is a less important, though statistically significant, predictor of organismal responses.'
acknowledgements: |

@@

## Relevant significance of variability today 
In addition to the importance of environmental variability across geologic timescales, organisms also interact with environmental variability on several shorter timescales. Organisms and the ecosystems they inhabit have experienced variation on many scales, from diurnal changes in light availability, to seasonal changes in temperature, to decadal changes in weather patterns (i.e. El Nino Southern Oscillation).  As such, the scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the tenacity and dynamics of ecosystems [khelifa_usefulness_2019; @stewart_mesocosm_2013]. How variability impacts populations may impact biodiversity, such that only certain species may survive, and how that biodiversity is maintained are key components to the full picture of ecosystem dynamics. As environmental patterns continue to change, from increased mean temperatures to more severe weather events (i.e. hurricanes), understanding the limits to organisms’ ability to cope with changes in environmental conditions is increasingly important for both individual level responses as well as how individual level impacts scale across population and community level responses [@helmuth_organismal_2010; @bernardo_biologically_2014; @huey_predicting_2012].

>>>>

---
title: 'Mean temperature, fluctuation magnitude, and level of biological organization affect biological responses amidst thermal variability: a meta analysis'
author: 'Margaret Anne Slein'
date: 'May 2021'
institution: 'Reed College'
division: 'Mathematics and Natural Sciences'
advisor: 'Sam Fey'
department: 'Biology'
degree: 'Bachelor of Arts'
header-includes:
    - \usepackage{setspace}\onehalfspacing
knit: bookdown::render_book
site: bookdown::bookdown_site
params:
  'Install needed packages for {thesisdown}': True
output:
  thesisdown::thesis_pdf: default 
abstract: 'Ecosystems and organisms have experienced variation on many temporal scales, from diurnal changes in light availability, to seasonal changes in temperature, to decadal changes in weather patterns. Though scientific literature has continued to highlight the importance of ecologically relevant studies for understanding the tenacity and dynamics of ecosystems, there has yet to be a quantitative review of how variability impacts biological responses across taxa and scales of biological organization. Here, we present a quantitative meta analysis of how thermal variability impacts biological responses across a variety of taxa, organism sizes, and biological organization. Our results suggest that the range of temperatures an organism experiences is the most important driver of response magnitude, with the interaction of mean temperature and fluctuation range emerging equally as significant predictors in our model. Further, our results also suggest that level of biological organization is a less important, though statistically significant, predictor of organismal responses.'
acknowledgements: |

@@

## Relevant significance of variability today 
In addition to the importance of environmental variability across geologic timescales, organisms also interact with environmental variability on several shorter timescales. Organisms and the ecosystems they inhabit have experienced variation on many scales, from diurnal changes in light availability, to seasonal changes in temperature, to decadal changes in weather patterns (i.e. El Niño Southern Oscillation).  As such, the scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the tenacity and dynamics of ecosystems [khelifa_usefulness_2019; @stewart_mesocosm_2013]. How variability impacts populations may impact biodiversity, such that only certain species may survive, and how that biodiversity is maintained are key components to the full picture of ecosystem dynamics. As environmental patterns continue to change, from increased mean temperatures to more severe weather events (i.e. hurricanes), understanding the limits to organisms’ ability to cope with changes in environmental conditions is increasingly important for both individual level responses as well as how individual level impacts scale across population and community level responses [@helmuth_organismal_2010; @bernardo_biologically_2014; @huey_predicting_2012].
